Basic Model formulation

### 2015-07-01 documentation/model_formulation.Rmd
update model formulation

```{r, echo=FALSE}

@@

workstation <-  list(host = "162.38.184.118", user = "schneider",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 10))

cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')

@@

```{r, echo=FALSE}
C <- function(rho, L = 5, g = 0.01) (g*rho*L)

@@

```{r, echo=FALSE}
C <- function(rho, L = 5, g = 0.01) g * L + rho*0

@@

```{r, echo=FALSE}
C <- function(rho, L = 5, a = 0.6, h = 100) (a*rho*L)/(1+a*h*rho)

@@

```{r, echo=FALSE}
C <- function(rho, L = 5, a = 0.6, h = 100, q = 1) (a*rho^(1+q)*L)/(1+a*h*(rho^(1+q)) )

@@

Given a parameter set  
```{r, echo=FALSE}

@@

```{r, echo=FALSE}
defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ), ylab = "growth/mortality" )

@@

```{r, echo=FALSE}
d_rho <- function(rho, parms, z = 4) { 
  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
}

@@

```{r, echo=FALSE}
par(mar = c(5,4,2,3)+0.1)
plot(rho,  with(parms,  m * rho + (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q))), ylab = "growth", type = "l", ylim = c(0,0.5))
  lines(rho, with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) ), col = "green3" )

@@

@@

```{r}

@@

```{r, echo=FALSE}
plot(I(rho_11/rho_1^2) ~ time , data = ODE_spex , type = "l", ylim = c(0,2))

@@

Visualising the attractor of the system is not as straightforward as for the mean-field models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria. In the spatially-explict pair-approximation model the attractor can be visualised if we assume that clustering in the initial state is non-random, and we assign starting values for $\rho_1$ and global mortality of the population $m_1$. To do this, we need to solve the mortality function towards $\rho{1\,,1}$ to obtain initial covers of all the different pairs and singletons.

Thus,

$$ m_1 = \rho_1 m_0 + \frac{(a+v\frac{\rho_{11}}{\rho_1})L(1-p\frac{\rho_{11}}{\rho_1}\rho_1^{1+q})}{1+(a+v\frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}} $$

rearranging yields:

$$ 0 = m_1 - m_0\rho_1+m_qah\rho_1^{1+q}-aL\rho_1^{1+q}-m_0ah\rho_1^{1+q} + (m_1v\rho_1^{q}+aLp\rho_1^{q}-vL\rho_1^{q}-m_0v\rho_1^{1+q})\rho_{11} + vLp\rho_1^{q-1}\rho_{11}^2$$
Polynomials of second order of the form $0 =  c+ bx +ax^2$ can be solved by the quadratic formula

$$x = \frac{-b\pm \sqrt{b^2-4ac} }{2a}$$

**apparently this solution is wrong!! Instead, I draw  initial $rho_11$ values systematically and elimitate those that produce invalid $m_ini$**

This works quite well, but produces unequal distributions of arrows along the y-axis.

```{r, echo=FALSE}
attractor(parms)
```

@@

```{r, echo=FALSE, fig.height = 3 }
par(mfrow = c(1,2))
parms$q = 0.4
bifurcation(parms, "L", c(0,12)) -> over_L

bifurcation(parms,"b", c(0,1)) -> over_b

@@

```{r, echo=FALSE, fig.height = 3}
par(mfrow = c(1,2))
bifurcation(parms, "L", pairapprox = TRUE, meanfield = FALSE, c(0,12)) -> over_L

bifurcation(parms,"b", pairapprox = TRUE, meanfield = FALSE, c(0,1)) -> over_b

@@

```{r, echo=FALSE, fig.height = 5}

@@

```{r, echo = FALSE}
over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
over_b$pairapprox$q_11[is.na(over_b$q_11)] <- 0

par(mfrow =c(1,2))
  plot(over_b$pairapprox$b, over_b$pairapprox$rho_1, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
```

@@

```{r, echo=FALSE, fig.width = 3}
par(mfrow = c(1,1))
  plot(over_b$pairapprox$rho_11, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
points(0,0, pch=20, col = "white", cex = 2)
 points(0,0, pch=20, cex = 0.5)

@@

```{r}

>>>>

```{r, echo=FALSE, warning=FALSE}

@@

workstation <-  list(host = "162.38.184.118", user = "florian",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 7))

cl <- makeSOCKcluster(workerlist,  outfile='out_messages.txt')

@@

```{r, echo=FALSE, warning=FALSE}
C <- function(rho, L = 5, g = 0.01) (g*rho*L)

@@

```{r, echo=FALSE, warning=FALSE}
C <- function(rho, L = 5, g = 0.01) g * L + rho*0

@@

```{r, echo=FALSE, warning=FALSE}
C <- function(rho, L = 5, a = 0.6, h = 100) (a*rho*L)/(1+a*h*rho)

@@

```{r, echo=FALSE, warning=FALSE}
C <- function(rho, L = 5, a = 0.6, h = 100, q = 1) (a*rho^(1+q)*L)/(1+a*h*(rho^(1+q)) )

@@

Given a parameter set  
```{r, echo=FALSE, warning=FALSE}

@@

```{r, echo=FALSE, warning=FALSE}
defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ), ylab = "growth/mortality" )

@@

```{r, echo=FALSE, warning=FALSE}
d_rho <- function(rho, parms, z = 4) { 
  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
}

@@

```{r, echo=FALSE, warning=FALSE}
par(mar = c(5,4,2,3)+0.1)
plot(rho,  with(parms,  m * rho + (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q))), ylab = "growth", type = "l", ylim = c(0,0.5))
  lines(rho, with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) ), col = "green3" )

@@

```{r, echo = FALSE, warning = FALSE, warning=FALSE}

@@

```{r, warning=FALSE}

@@

```{r, echo=FALSE, warning=FALSE}
plot(I(rho_11/rho_1^2) ~ time , data = ODE_spex , type = "l", ylim = c(0,2))

@@

Visualising the attractor of the system is not as straightforward as for the mean-field models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria. In the spatially-explict pair-approximation model the attractor can be visualised if we assume that clustering in the initial state is non-random, and we assign starting values for $\rho_1$ and global mortality of the population $m_1$.

I assign inital values of mortality equally along the x-axis for the assumption of highest and lowest clustering. The development of global and local cover yields timeseries of average growth and mortality for all those virtualised cases which can be plotted as arrowstowards the attractors.

```{r, echo=FALSE, warning=FALSE}
attractor(parms)
```

This nicely illustrates that the unstable equilibrium now turns from a simple threshold point to a line or surface.

@@

```{r, echo=FALSE, fig.height = 3, warning=FALSE }
par(mfrow = c(1,2))
parms$q = 0.4
bifurcation(parms, "L", c(0,12), returnoutput = TRUE) -> over_L

bifurcation(parms,"b", c(0,1), returnoutput = TRUE) -> over_b

@@

```{r, echo=FALSE, fig.height = 3, warning=FALSE}
par(mfrow = c(1,2))
bifurcation(parms, "L", pairapprox = TRUE, meanfield = FALSE, c(0,12), returnoutput = TRUE) -> over_L

bifurcation(parms,"b", pairapprox = TRUE, meanfield = FALSE, c(0,1), returnoutput = TRUE) -> over_b

@@

```{r, echo=FALSE, fig.height = 5, warning=FALSE}

@@

```{r, echo = FALSE, warning=FALSE}
over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
over_b$pairapprox$q_11[is.na(over_b$q_11)] <- 0

par(mfrow =c(1,2))
  plot(over_b$pairapprox$b, over_b$pairapprox$rho_1, 
       pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1),
       xlab = "environmental quality", ylab = "global cover"
       )
  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, 
       pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1),
       xlab = "environmental quality", ylab = "local cover")
```

@@

```{r, echo=FALSE, fig.width = 3, fig.height = 3, warning=FALSE}
par(mfrow = c(1,1), mar = c(4,4,1,1))
  plot(over_b$pairapprox$rho_1, over_b$pairapprox$q_11, 
       pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1),
       xlab = "global cover", ylab = "local cover")
abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
points(0,0, pch=20, col = "white", cex = 2)
 points(0,0, pch=20, cex = 0.5)

@@

```{r, echo=FALSE, warning=FALSE}
